---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# trrrj

<!-- badges: start -->
[![R build status](https://github.com/euctrl-pru/trrrj/workflows/Render README/badge.svg)](https://github.com/euctrl-pru/trrrj/actions)
[![Travis build status](https://travis-ci.org/euctrl-pru/trrrj.svg?branch=master)](https://travis-ci.org/euctrl-pru/trrrj)
[![codecov](https://codecov.io/gh/euctrl-pru/trrrj/branch/master/graph/badge.svg)](https://codecov.io/gh/euctrl-pru/trrrj)
<!-- badges: end -->



The `{trrrj}` package provides facilities to aid in the analysis of flight trajectories.

It is in contiuous flux both because new features are added but also because of users'
feedback.

If you want to contribute but you feel overwhelmed by the git/github scaffolding, please have a 
look at the `usethis` vignette "[Pull request helpers][prh]". It marvellously describes
a nice workflow and would make collaboration very easy and fruitful.

[prh]: https://usethis.r-lib.org/articles/articles/pr-functions.html


## Installation

`{trrrj}` is not yet available from CRAN, but you can install the development
version from github with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("euctrl-pru/trrrj")
```

## Usage

There are functions that allow you to:

* Load trajectories from various providers/sources:
    - archived/live feed [Flightradar24][fr24] ADS-B based (files)
    - live feed [Flightradar24][fr24] ADS-B based (EUROCONTROL's DB)
    - archived [Network Manager][nm-cfmu] CPR's (files); see [CPRs] note
    - [DDR2][ddr2] SO6 trajectories (EUROCONTROL's DB)
    - historical data from [OpenSky Network][osn]

* Plot trajectories
    - 2D plot
    - vertical profiles (time or distance based)
    
* Analyse trajectories (Under work: more to come!)
    - extract level flight segments

* Read/transform/*save your day*
    - parse Longitude/Latitude in various formats
    - determine the AIRAC cyle for a date or its range (ICAO or CFMU format); see [AIRAC] note


## Data

The package contains small data source files in order to
provide realistic examples and use cases for guides and vignettes.



## Tutorials and Guides et al.

The vignette [Plotting trajectories](https://trrrj.ansperformance.eu/articles/trrrj-plotting.html)
provides a good introduction on how to use the package.

The [`{osn}` package](https://github.com/espinielli/osn) can be used to access
the [OpenSky Network Data](osn).


## Notes

### CPRs
Correlated Position Reports (CPR) are position report provided by the ATC radar facilities.
They are *correlated* because they contain flight information

### AIRAC
An AIRAC (Aeronautical Information Regulation And Control) cycle is the 28-day period
that regulates the [pubblication of aeronautical information][airac]


[airac]: <https://en.wikipedia.org/wiki/Aeronautical_Information_Publication> "AIP - Aeronautical Information Publication"
[fr24]: <https://www.flightradar24.com/> "Flightradar24"
[ddr2]: <https://www.eurocontrol.int/ddr> "Demand Data Repository - EUROCONTROL"
[nm-cfmu]: <https://www.eurocontrol.int/network-manager> "Network Manager - EUROCONTROL"
[cpr]: <https://ansperformance.eu/acronym/cpr/> "Correlated Position Report"
[osn]: <https://opensky-network.org/> "OpenSky Network"
