---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# trrrj

[![Travis build status](https://travis-ci.org/euctrl-pru/trrrj.svg?branch=master)](https://travis-ci.org/euctrl-pru/trrrj)
[![codecov](https://codecov.io/gh/euctrl-pru/trrrj/branch/master/graph/badge.svg)](https://codecov.io/gh/euctrl-pru/trrrj)



The `{trrrj}` package provides facilities to aid in the analysis of flight trajectories.

## Installation

`{trrrj}` is not yet available from CRAN, but you can install the development
version from github with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("euctrl-pru/trrrj")
```

## Usage

There are functions that allow you to:

* Load trajectories from various providers/sources:
    - archived/live feed [Flightradar24][fr24] ADS-B based (files)
    - live feed [Flightradar24][fr24] ADS-B based (EUROCONTROL's DB)
    - archived [Network Manager][nm-cfmu] CPR's^[Correlated Position Reports (CPR)
      are position report provided by the ATC radar facilities.
      They are *correlated* because they contain flight information] (files)
    - [DDR2][ddr2] SO6 trajectories (EUROCONTROL's DB)

* Plot trajectories
    - 2D plot
    - vertical profiles (time or distance based)
    
* Analyse trajectories (Under work: more to come!)
    - extract level flight segments

* Read/transform/*save your day*
    - parse Longitude/Latitude in various formats
    - determine the AIRAC^[An AIRAC (Aeronautical Information Regulation And Control) cycle
      is the 28-day period that regulates the [pubblication of aeronautical information](https://en.wikipedia.org/wiki/Aeronautical_Information_Publication)] cyle
      for a date or its range (ICAO or CFMU format)


## Data

The package contains small data source files in order to
provide realistic examples and use cases for guides and vignettes.



## Tutorials and Guides

The vignette [Plotting trajectories](articles/trrrj-plotting.html) provides
a good introduction on how to use the package.


[fr24]: <https://www.flightradar24.com/> "Flightradar24"
[ddr2]: <https://www.eurocontrol.int/ddr> "Demand Data Repository - EUROCONTROL"
[nm-cfmu]: <https://www.eurocontrol.int/network-manager> "Network Manager - EUROCONTROL"
[cpr]: <https://ansperformance.eu/acronym/cpr/> "Correlated Position Report"
